---
title: "One Cell Type"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{One Cell Type}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SplitClusterTest)
```

The simulated data `simdata_1ct` with one cell type (i.e., no cluster structure) is generated based on the real single-cell data `DuoClustering2018::sce_full_Zhengmix4eq()` with the help of `scDesign3`. For more details about generating synthetic data, please check our paper.

The structure of `simdata_1ct` is as follows:

```{r}
str(simdata_1ct)
```

Since there is no cluster structure, so the index of DE genes is empty:

```{r}
simdata_1ct$de_idx
```

Our proposed multiple data splitting (MDS) does not return significant DE genes:

```{r}
mss = mds1(simdata_1ct$simu_sce, M = 1, 
           params1 = list(normalized_method = "sct", pca.whiten = TRUE),
           params2 = list(normalized_method = "sct", pca.whiten = TRUE))
sel = mds2(mss)
```

The mirror statistics are distributed as follows:

```{r hist-mirror-one, fig.alt="Histogram of Mirror Statistics (One Cell Type)"}
hist(mss[[1]], breaks = 50)
```

However, the naive double-dipping method will return many false positives:

```{r}
sel.dd = dd(simdata_1ct$simu_sce, params = list(normalized_method = "sct", pca.whiten = TRUE))
length(sel.dd)
```

## Session Info

```{r}
sessionInfo()
```
