---
title: "Two Gaussians"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two Gaussians}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SplitClusterTest)
```

Imagine you have data from two distinct clusters, each with 200 features and 50 observations. The interesting part: only 10% of these features (20 features) actually differ between the groups. The remaining features are identically distributed. The magnitude of these differences is controlled by a parameter called `delta`.

## Generating Example Data

Let's create such a dataset using the `gen_data_normal` function:

```{r}
data = gen_data_normal(n = 100, p = 200, prop = 0.1, delta = 3)
X = data$X
L = data$L
```

In this example:

- Features 1 through 20 are truly different between clusters (`delta = 3`)
- Features 21 through 200 are identically distributed (no difference)

## Identifying Non-null Features by DS

Now, let's use our DS procedure to automatically identify which features distinguish the two clusters while controlling the false discovery rate (FDR) at a nominal level of `q = 0.2`:

```{r}
set.seed(1)
res = ds(X, q = 0.2)
names(res$sel_set)
```

## Evaluating Performance

We know the ground truth: features 1–20 are truly different. Let's measure how accurately DS identified them:

```{r}
calc_acc(res$sel_set, 1:20)
```

This gives us:

- **Precision (1 - FDR)**: Proportion of selected features that are truly different
- **Recall (Power)**: Proportion of true features that were correctly identified
- **F1 score**: Harmonic mean of precision and recall


## Visualizing Mirror Statistics

DS uses mirror statistics to distinguish signal from noise. Let's visualize their distribution:

```{r}
hist(res$ms)
```

**Interpretation**: Mirror statistics with large positive values suggest strong evidence of differences between clusters.


## A More Robust Approach: MDS

For increased robustness against variability, you can use the MDS (Multiple DS) procedure, which aggregates results across multiple random splits:

```{r}
set.seed(1)
res = mds(X, M = 10, q = 0.2)
```

Check the accuracy of the selected set

```{r}
calc_acc(res, 1:20)
```

## 💡 Key Takeaways

- DS provides a straightforward approach to feature selection with FDR control
- MDS offers enhanced robustness through aggregation across multiple random splits
- Both methods aim to identify the 20 truly different features while controlling false discoveries
- The histogram of mirror statistics helps visualize the separation between signal and noise


